---
title: "Repeability of Frog Morphometric Measurements"
author:
  - name: Marguerite Butler
    url: https://butlerlab.org
    affiliation: School of Life Sciences, University of Hawaii
    affiliation_url: https://manoa.hawaii.edu/lifesciences/
description: "Within and between measurer repeatability for the ecomorph analysis"
date: "`r format(Sys.Date())`"
toc: true
format: 
  html: 
    embed-resources: true
  pdf: default
---

# Assessing Repeatability of Morphometric Measurements 

We want to measure repeatability of our morphometric measurements, both within individual measurers and between measurers. We plan to combine data with that measured by Julio Rivera in 2015, so we want to know that the data are compatible. 

The accuracy we are aiming for is 95%, or 5% measurement error (both within and between individual measurers). 

Each measurer measured the same 7 frogs (_Hylophorbus sp._ from Buyetai: JR306,308,311,320,321,324), which was repeated (2 sets of measurements per measurer). 

### Setup 

```{r}
#| warning: false
require( googlesheets4 )
require(dplyr)
require(magrittr)
require(ggplot2)
require(reshape2)
require(cowplot)

if(dir.exists("output")!=TRUE) dir.create("output") # check if output directory out exists, if false create
```

Read in our [data from a Google Spreadsheet](https://docs.google.com/spreadsheets/d/1-w62GXvKwQ868dwiMVzgPdLYbgSkXxSqasE-YoaGOUc/edit#gid=0), and convert measurer, session and jr number to factors.

```{r}
file <- "https://docs.google.com/spreadsheets/d/1-w62GXvKwQ868dwiMVzgPdLYbgSkXxSqasE-YoaGOUc/edit#gid=0"

gs4_deauth() # not a private sheet, so no need for authentication
alldat <- as.data.frame(read_sheet(file)) %>%  ### INPUT DATA from googlesheet 
         filter(!is.na(measurer))			# remove spacer rows

alldat %<>% mutate_at(c("measurer", "session", "jr"), as.factor) %>%
    mutate( shape = case_when( measurer=="JR" ~ 25, measurer!="JR" ~ 19 )) %>%
    mutate( shape = as.factor(shape))

jrblock <- alldat[alldat$measurer=="JR",]  # copy JR data into all sessions (0,1,2)
jrblock3 <- jrblock2 <- jrblock0 <- jrblock
jrblock0$session <- 0    
jrblock2$session <- 2
jrblock3$session <- 3
alldat <- rbind(alldat, jrblock0, jrblock2, jrblock3)
```

## Repeatability 

This requires two full sets of measurements, so this is still in progress. 

```{r}
mod <- with(alldat, summary(aov(lm( svl ~ jr ))))

s2_within <- ms_within <- mod[[1]][2,3]
s2_within

ms_among <- mod[[1]][1,3]
s2_among <- (ms_among-ms_within)/2
ME <- s2_within/(s2_within+s2_among) * 100
ME

```

# Functions 

These function creates the plots that are repeated for each morphometric variable.

```{r}
dat <- alldat %>% filter(session==0)

p <- dat %>% ggplot(aes(svl, femur, color=measurer, label=jr))
v <- dat %>% ggplot(aes(jr, svl, group=jr, color=measurer, shape=measurer))


make_plots <- function(p, v) {
  q1 <- p + geom_point(size = 3) + 
	geom_smooth( aes(group=measurer), method="lm", alpha=.1) + 	
	geom_text(nudge_y = .15) +
	theme_bw()
	
  q2 <- p + geom_point(size = 3) + 
	geom_smooth( method="lm", alpha=.1) + 	
	geom_text(nudge_y = 1) +
	facet_grid( measurer ~ . ) +
	theme_bw()

   plot_grid(q1, q2, labels="AUTO")	

 }

make_violins <- function(v) {

 v + geom_violin() + 
  geom_jitter(aes(x=jr, color=measurer, shape=measurer, size=measurer), width=.2) + 
  scale_shape_manual(values=c(19,19,17,19,19,19)) +
  scale_size_manual(values=c(3,3,5,3,3,3))

}
```
Set up the ggplots: 

```{r}
dat <- alldat %>% filter(session==0)  # filter just for session 0 (naive)
 
femp <- dat %>% ggplot(aes(svl, femur, color=measurer, label=jr))
tibp <- dat %>% ggplot(aes(svl, tibiofibula, color=measurer, label=jr))
tarp <- dat %>% ggplot(aes(svl, tarsus, color=measurer, label=jr))
footp <- dat %>% ggplot(aes(svl, foot, color=measurer, label=jr))
hwp <- dat %>% ggplot(aes(svl, headW, color=measurer, label=jr))
hlp <- dat %>% ggplot(aes(svl, headL, color=measurer, label=jr))
hump <- dat %>% ggplot(aes(svl, humerus, color=measurer, label=jr))
radp <- dat %>% ggplot(aes(svl, radioulna, color=measurer, label=jr))
handp <- dat %>% ggplot(aes(svl, hand, color=measurer, label=jr))
  
sv <- dat %>% ggplot(aes(jr, svl, group=jr))
femv <- dat %>% ggplot(aes(jr, femur, group=jr))
tibv <- dat %>% ggplot(aes(jr, tibiofibula, group=jr))
tarv <- dat %>% ggplot(aes(jr, tarsus, group=jr))
footv <- dat %>% ggplot(aes(jr, foot, group=jr))
hwv <- dat %>% ggplot(aes(jr, headW, group=jr))
hlv <- dat %>% ggplot(aes(jr, headL, group=jr))
humv <- dat %>% ggplot(aes(jr, humerus, group=jr))
radv <- dat %>% ggplot(aes(jr, radioulna, group=jr))
handv <- dat %>% ggplot(aes(jr, hand, group=jr))
```

# Validating Landmarks (session 0)

Plot all of our data against the original dataset (JR in green), as well as faceted by individual. The goal here is to validate the landmarks.  

### SVL

```{r}
#| warning: false
make_violins(sv)
```
Looks like everyoneʻs SVL are a little short. Basically JR304 and JR306 match well, but the others, JR308, JR311, JR320, JR321, JR322, JR324 are a little short. Did he press them down? Or use a ruler?  

Diana and Ke investigated and found that __gently pressing on the frog__ was necessary to reapeat JRʻs measurements. Just enough to maximize the length. 

### Femur

```{r}
#| warning: false
make_violins(femv)
```
```{r}
#| warning: false
make_plots( femp )
```
Looks like my femur measurements (MB) are a little short - JR must have measured from the vent rather than from the midline perpendicular to the femur. 

The landmarks are from __the edge of the knee to the midline at the vent__. 

### Tibiofibula

```{r}
#| warning: false
make_violins(tibv)
```
```{r}
#| warning: false
make_plots( tibp )
```
Tibiofibula looks good, except for what looks like a typo in DGʻs data?

We measured the bone, from end to end.

### Tarsus

```{r}
#| warning: false
make_violins(tarv)
```
```{r}
#| warning: false
make_plots( tarp )
```
Not sure what I did! Lol. Good example of the unfortunate outlier being the largest throwing the whole trend off. 
A few measurers are off on some specimens.

### Foot

```{r}
#| warning: false
make_violins(footv)
```
```{r}
#| warning: false
make_plots( footp )
```
Looks good! 

### Head Width

```{r}
#| warning: false
make_violins(hwv)
```
```{r}
#| warning: false
make_plots( hwp )
```

Looks good! Measured at the widest part of the head or at the center of the tympanum (no further back than the tympanum). 

### Head Length

```{r}
#| warning: false
make_violins(hlv)
```
```{r}
#| warning: false
make_plots( hlp )
```

Oh oh. We are not using the right Head Length landmark.

Diana and Ke discovered the the measurement is from the tip of the snout to the center of the tympanum.

### Humerus

```{r}
#| warning: false
make_violins(humv)
```
```{r}
#| warning: false
make_plots( hump )
```

Hmm. We are all a little consistently lower than JR, some more so.

The landmark for the end of the humerus can be most easily found by gently pressing the caliper onto the chest, which will move the arm. It is from __the articulation of the humerus with the scapula to the edge of the elbow__. 

### Radioulna

```{r}
#| warning: false
make_violins(radv)
```
```{r}
#| warning: false
make_plots( radp)
```

Need to work on this one too.

### Hand
```{r}
#| warning: false
make_violins(handv)
```
```{r}
#| warning: false
make_plots( handp )
```
Looks good!

## Conclusions

We need to confirm the landmarks JR used for the limb segment and head length measurements (foot and hand are OK). It looks promising for repeatability but we need to confirm after we get two full sets of measurements. 

# Session 1

Set up the ggplots: 

```{r}
dat <- alldat %>% filter(session==1)

femp <- dat %>% ggplot(aes(svl, femur, color=measurer, label=jr))
tibp <- dat %>% ggplot(aes(svl, tibiofibula, color=measurer, label=jr))
tarp <- dat %>% ggplot(aes(svl, tarsus, color=measurer, label=jr))
footp <- dat %>% ggplot(aes(svl, foot, color=measurer, label=jr))
hwp <- dat %>% ggplot(aes(svl, headW, color=measurer, label=jr))
hlp <- dat %>% ggplot(aes(svl, headL, color=measurer, label=jr))
hump <- dat %>% ggplot(aes(svl, humerus, color=measurer, label=jr))
radp <- dat %>% ggplot(aes(svl, radioulna, color=measurer, label=jr))
handp <- dat %>% ggplot(aes(svl, hand, color=measurer, label=jr))

sv <- dat %>% ggplot(aes(jr, svl, group=jr))
femv <- dat %>% ggplot(aes(jr, femur, group=jr))
tibv <- dat %>% ggplot(aes(jr, tibiofibula, group=jr))
tarv <- dat %>% ggplot(aes(jr, tarsus, group=jr))
footv <- dat %>% ggplot(aes(jr, foot, group=jr))
hwv <- dat %>% ggplot(aes(jr, headW, group=jr))
hlv <- dat %>% ggplot(aes(jr, headL, group=jr))
humv <- dat %>% ggplot(aes(jr, humerus, group=jr))
radv <- dat %>% ggplot(aes(jr, radioulna, group=jr))
handv <- dat %>% ggplot(aes(jr, hand, group=jr))
```

### SVL

```{r}
#| warning: false
make_violins(sv)
```
Much better!

### Femur

```{r}
#| warning: false
make_violins(femv)
```
```{r}
#| warning: false
make_plots( femp )
```
Better!

### Tibiofibula

```{r}
#| warning: false
make_violins(tibv)
```
```{r}
#| warning: false
make_plots( tibp )
```
Better!

### Tarsus

```{r}
#| warning: false
make_violins(tarv)
```
```{r}
#| warning: false
make_plots( tarp )
```
Better but Diana still a bit low and Ke a bit high

### Foot

```{r}
#| warning: false
make_violins(footv)
```
```{r}
#| warning: false
make_plots( footp )
```
Still good! 

### Head Width

```{r}
#| warning: false
make_violins(hwv)
```
```{r}
#| warning: false
make_plots( hwp )
```

Good. I may have been too generous.

### Head Length

```{r}
#| warning: false
make_violins(hlv)
```
```{r}
#| warning: false
make_plots( hlp )
```

Much better! 

### Humerus

```{r}
#| warning: false
make_violins(humv)
```
```{r}
#| warning: false
make_plots( hump )
```

Several folks are still a bit low. 

### Radioulna

```{r}
#| warning: false
make_violins(radv)
```
```{r}
#| warning: false
make_plots( radp)
```

Now looks like random error, still a bit high though.

### Hand
```{r}
#| warning: false
make_violins(handv)
```
```{r}
#| warning: false
make_plots( handp )
```

Now looks like random error.

# Session 2


```{r}
dat <- alldat %>% filter(session==2)

femp <- dat %>% ggplot(aes(svl, femur, color=measurer, label=jr))
tibp <- dat %>% ggplot(aes(svl, tibiofibula, color=measurer, label=jr))
tarp <- dat %>% ggplot(aes(svl, tarsus, color=measurer, label=jr))
footp <- dat %>% ggplot(aes(svl, foot, color=measurer, label=jr))
hwp <- dat %>% ggplot(aes(svl, headW, color=measurer, label=jr))
hlp <- dat %>% ggplot(aes(svl, headL, color=measurer, label=jr))
hump <- dat %>% ggplot(aes(svl, humerus, color=measurer, label=jr))
radp <- dat %>% ggplot(aes(svl, radioulna, color=measurer, label=jr))
handp <- dat %>% ggplot(aes(svl, hand, color=measurer, label=jr))

sv <- dat %>% ggplot(aes(jr, svl, group=jr))
femv <- dat %>% ggplot(aes(jr, femur, group=jr))
tibv <- dat %>% ggplot(aes(jr, tibiofibula, group=jr))
tarv <- dat %>% ggplot(aes(jr, tarsus, group=jr))
footv <- dat %>% ggplot(aes(jr, foot, group=jr))
hwv <- dat %>% ggplot(aes(jr, headW, group=jr))
hlv <- dat %>% ggplot(aes(jr, headL, group=jr))
humv <- dat %>% ggplot(aes(jr, humerus, group=jr))
radv <- dat %>% ggplot(aes(jr, radioulna, group=jr))
handv <- dat %>% ggplot(aes(jr, hand, group=jr))
```

### SVL

```{r}
#| warning: false
make_violins(sv)
```
Much better!

### Femur

```{r}
#| warning: false
make_violins(femv)
```
```{r}
#| warning: false
make_plots( femp )
```
Better!

### Tibiofibula

```{r}
#| warning: false
make_violins(tibv)
```
```{r}
#| warning: false
make_plots( tibp )
```
Better!

### Tarsus

```{r}
#| warning: false
make_violins(tarv)
```
```{r}
#| warning: false
make_plots( tarp )
```
Better but Diana still a bit low and Ke a bit high

### Foot

```{r}
#| warning: false
make_violins(footv)
```
```{r}
#| warning: false
make_plots( footp )
```
Still good! 

### Head Width

```{r}
#| warning: false
make_violins(hwv)
```
```{r}
#| warning: false
make_plots( hwp )
```

Good. I may have been too generous.

### Head Length

```{r}
#| warning: false
make_violins(hlv)
```
```{r}
#| warning: false
make_plots( hlp )
```

Much better! 

### Humerus

```{r}
#| warning: false
make_violins(humv)
```
```{r}
#| warning: false
make_plots( hump )
```

Several folks are still a bit low. 

### Radioulna

```{r}
#| warning: false
make_violins(radv)
```
```{r}
#| warning: false
make_plots( radp )
```

Now looks like random error, still a bit high though.

### Hand
```{r}
#| warning: false
make_violins(handv)
```
```{r}
#| warning: false
make_plots( handp )
```

# All together 

```{r}
dat <- alldat 

sv <- dat %>% ggplot(aes(jr, svl, group=jr))
femv <- dat %>% ggplot(aes(jr, femur, group=jr))
tibv <- dat %>% ggplot(aes(jr, tibiofibula, group=jr))
tarv <- dat %>% ggplot(aes(jr, tarsus, group=jr))
footv <- dat %>% ggplot(aes(jr, foot, group=jr))
hwv <- dat %>% ggplot(aes(jr, headW, group=jr))
hlv <- dat %>% ggplot(aes(jr, headL, group=jr))
humv <- dat %>% ggplot(aes(jr, humerus, group=jr))
radv <- dat %>% ggplot(aes(jr, radioulna, group=jr))
handv <- dat %>% ggplot(aes(jr, hand, group=jr))

wrap_violin <- function(v) { 
  v + geom_violin() +
    geom_jitter(aes(x=jr, color=measurer, shape=measurer, size=measurer), width=.2) + 
    scale_shape_manual(values=c(19,19,17,19,19,19)) +
    scale_size_manual(values=c(3,3,5,3,3,3)) +
    facet_wrap(. ~ session )  
}
```

### SVL

```{r}
#| warning: false
wrap_violin(sv)
```

### Femur
```{r}
#| warning: false
wrap_violin(femv)
```
### Tibiofibula
```{r}
#| warning: false
wrap_violin(tibv)
```
### Tarsus
```{r}
#| warning: false
wrap_violin(tarv)
```

For round 3, I (MB) measured only the tarsus bone length and not the tarsus segment length. It seems JR measured the limb segment. 

### Foot
```{r}
#| warning: false
wrap_violin(footv)
```

### Head Width
```{r}
#| warning: false
wrap_violin(hwv)
```

### Head Length
```{r}
#| warning: false
wrap_violin(hlv)
```

### Humerus
```{r}
#| warning: false
wrap_violin(humv)
```

### Radioulna
```{r}
#| warning: false
wrap_violin(radv)
```

### Hand
```{r}
#| warning: false
wrap_violin(handv)
```

