---
title: "Repeability of Frog Morphometric Measurements"
author:
  - name: Marguerite Butler
    url: https://butlerlab.org
    affiliation: School of Life Sciences, University of Hawaii
    affiliation_url: https://manoa.hawaii.edu/lifesciences/
description: "Within and between measurer repeatability for the ecomorph analysis"
date: "`r format(Sys.Date())`"
format: 
  html: default
  pdf: default
---

# Assessing Repeatability of Morphometric Measurements 

We want to measure repeatability of our morphometric measurements, both within individual measurers and between measurers. We plan to combine data with that measured by Julio Rivera in 2015, so we want to know that the data are compatible. 

The accuracy we are aiming for is 95%, or 5% measurement error (both within and between individual measurers). 

Each measurer measured the same 7 frogs (_Hylophorbus sp._ from Buyetai: JR306,308,311,320,321,324), which was repeated (2 sets of measurements per measurer). 

### Setup 

```{r}
#| warning: false
require( googlesheets4 )
require(dplyr)
require(magrittr)
require(ggplot2)
require(cowplot)

if(dir.exists("output")!=TRUE) dir.create("output") # check if output directory out exists, if false create
```

Read in our [data from a Google Spreadsheet](https://docs.google.com/spreadsheets/d/1-w62GXvKwQ868dwiMVzgPdLYbgSkXxSqasE-YoaGOUc/edit#gid=0), and convert measurer, session and jr number to factors.

```{r}
file <- "https://docs.google.com/spreadsheets/d/1-w62GXvKwQ868dwiMVzgPdLYbgSkXxSqasE-YoaGOUc/edit#gid=0"

gs4_deauth() # not a private sheet, so no need for authentication
dat <- as.data.frame(read_sheet(file)) %>%  ### INPUT DATA from googlesheet 
         filter(!is.na(measurer))			# remove spacer rows

dat %<>% mutate_at(c("measurer", "session", "jr"), as.factor) %>%
    mutate( shape = case_when( measurer=="JR" ~ 25, measurer!="JR" ~ 19 )) %>%
    mutate( shape = as.factor(shape))

```

## Repeatability 

This requires two full sets of measurements, so this is still in progress. 

```{r}
mod <- with(dat, summary(aov(lm( svl ~ jr ))))

s2_within <- ms_within <- mod[[1]][2,3]
s2_within

ms_among <- mod[[1]][1,3]
s2_among <- (ms_among-ms_within)/2
ME <- s2_within/(s2_within+s2_among) * 100
ME
```

## Plots of measurements from the first session

Plot all of our data against the original dataset (JR in green), as well as faceted by individual.

This function creates the plots that are repeated for each morphometric variable.

```{r}
p <- dat %>% ggplot(aes(svl, femur, color=measurer, label=jr))
v <- dat %>% ggplot(aes(jr, svl, group=jr, color=measurer, shape=measurer))

make_violins <- function(v) {

 v + geom_violin() + 
  geom_jitter(aes(x=jr, color=measurer, shape=measurer, size=measurer), width=.2) + 
  scale_shape_manual(values=c(19,19,17,19,19,19)) +
  scale_size_manual(values=c(3,3,5,3,3,3))

}

make_plots <- function(p, v) {
  q1 <- p + geom_point(size = 3) + 
	geom_smooth( aes(group=measurer), method="lm", alpha=.1) + 	
	geom_text(nudge_y = .15) +
	theme_bw()
	
  q2 <- p + geom_point(size = 3) + 
	geom_smooth( method="lm", alpha=.1) + 	
	geom_text(nudge_y = 1) +
	facet_grid( measurer ~ . ) +
	theme_bw()

   plot_grid(q1, q2, labels="AUTO")	

 }

# q3 <- v + geom_violin() + 
#  geom_jitter(aes(x=jr, color=measurer, shape=measurer), size=3, width=.2) + 
#  scale_shape_manual(values=c(19,19,17,19,19,19)) 

# q3
```
Set up the ggplots: 

```{r}
femp <- dat %>% ggplot(aes(svl, femur, color=measurer, label=jr))
tibp <- dat %>% ggplot(aes(svl, tibiofibula, color=measurer, label=jr))
tarp <- dat %>% ggplot(aes(svl, tarsus, color=measurer, label=jr))
footp <- dat %>% ggplot(aes(svl, foot, color=measurer, label=jr))
hwp <- dat %>% ggplot(aes(svl, headW, color=measurer, label=jr))
hlp <- dat %>% ggplot(aes(svl, headL, color=measurer, label=jr))
hump <- dat %>% ggplot(aes(svl, humerus, color=measurer, label=jr))
radp <- dat %>% ggplot(aes(svl, radioulna, color=measurer, label=jr))
handp <- dat %>% ggplot(aes(svl, hand, color=measurer, label=jr))

sv <- dat %>% ggplot(aes(jr, svl, group=jr, color=measurer, shape=measurer))
femv <- dat %>% ggplot(aes(jr, femur, group=jr, color=measurer, shape=measurer))
tibv <- dat %>% ggplot(aes(jr, tibiofibula, group=jr, color=measurer, shape=measurer))
tarv <- dat %>% ggplot(aes(jr, tarsus, group=jr, color=measurer, shape=measurer))
footv <- dat %>% ggplot(aes(jr, foot, group=jr, color=measurer, shape=measurer))
hwv <- dat %>% ggplot(aes(jr, headW, group=jr, color=measurer, shape=measurer))
hlv <- dat %>% ggplot(aes(jr, headL, group=jr, color=measurer, shape=measurer))
humv <- dat %>% ggplot(aes(jr, humerus, group=jr, color=measurer, shape=measurer))
radv <- dat %>% ggplot(aes(jr, radioulna, group=jr, color=measurer, shape=measurer))
handv <- dat %>% ggplot(aes(jr, hand, group=jr, color=measurer, shape=measurer))
```

Make plots:

### SVL

```{r}
#| warning: false
make_violins(sv)
```
Looks like everyoneʻs SVL are a little short. Basically JR304 and JR306 match well, but the others, JR308, JR311, JR320, JR321, JR322, JR324 are a little short. Did he press them down? Or use a ruler?  


### Femur

```{r}
#| warning: false
make_violins(femv)
```
```{r}
#| warning: false
make_plots( femp )
```
Looks like my femur measurements (MB) are a little short - JR must have measured from the vent rather than from the midline perpendicular to the femur. 

### Tibiofibula

```{r}
#| warning: false
make_violins(tibv)
```
```{r}
#| warning: false
make_plots( tibp )
```
Tibiofibula looks good, except for what looks like a typo in DGʻs data?

### Tarsus

```{r}
#| warning: false
make_violins(tarv)
```
```{r}
#| warning: false
make_plots( tarp )
```
Not sure what I did! Lol. Good example of the unfortunate outlier being the largest throwing the whole trend off. 

### Foot

```{r}
#| warning: false
make_violins(footv)
```
```{r}
#| warning: false
make_plots( footp )
```
Looks good! 

### Head Width

```{r}
#| warning: false
make_violins(hwv)
```
```{r}
#| warning: false
make_plots( hwp )
```

Looks good!

### Head Length

```{r}
#| warning: false
make_violins(hlv)
```
```{r}
#| warning: false
make_plots( hlp )
```

Oh oh. We are not using the right Head Length landmark.

### Humerus

```{r}
#| warning: false
make_violins(humv)
```
```{r}
#| warning: false
make_plots( hump )
```

Hmm. We are all a little consistently lower than JR, some more so.

### Radioulna

```{r}
#| warning: false
make_violins(radv)
```
```{r}
#| warning: false
make_plots( radp)
```

Need to work on this one too.

### Hand
```{r}
#| warning: false
make_violins(handv)
```
```{r}
#| warning: false
make_plots( handp )
```
Looks good!

## Conclusions

We need to confirm the landmarks JR used for the limb segment and head length measurements (foot and hand are OK). It looks promising for repeatability but we need to confirm after we get two full sets of measurements. 
